API design for machine learning software: experiences from the scikit-learn project

### 2013-06-26 paper.tex
Improve GUI / command line comparison.

\section{Comparison with other packages}

\label{sec:comparison}

\textcolor{red}{TODO: emphasize scikit-learn's place in the overall scipy ecosystem, and its role in spurring the development of more basic tools.}\\

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.

Many machine learning packages exist, with new ones still often appearing,
that are intended to be used as command-line tools
(and sometimes do not offer any type of API).
Examples include Weka \citep{hall2009weka}, SofiaML \citep{sculley2010web}
and Vowpal~Wabbit \citep{langford2009sparse}.
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need to do programming,
and will do so in a variety of languages.
We argue that this makes it harder for researchers and developers
to exchange software and collaborate on software.

>>>>

\section{Related software}
\label{sec:comparison}

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). We argue that a GUI
lacks flexibility, makes experiments difficult to reproduce and has a high
maintenance cost from a development point of view. For these reasons, we make no
attempt to provide such a GUI in \sklearn.
Other existing machine learning packages 
such as SofiaML \citep{sculley2010web}
and Vowpal~Wabbit \citep{langford2009sparse}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process inputs/outputs,
and will do so in a variety of languages.
We argue that this makes it harder for researchers and developers
to exchange and collaborate on software.
